Abstractions for Multi-Sorted Substitutions

### 2024-03-16 paper/paper.tex
Update.

  \AppCode\FPreserve
  The type of \ACode{\ADef{\_⊢⋯\_}} is the same as in the record definition and hence omitted.
  The interesting parts of the proof are:
  \begin{itemize}
  \item
    There is a strong similarity to the instantiation of
    map traversal \ACode{\AField{\_⋯\_}}. Where \ACode{\AField{\_⋯\_}}
    used \ACode{\AField{`/id}} or \ACode{\AField{\_↑\_}}, we now use their
    preservation lemmas \ACode{\AField{⊢`/id}} or \ACode{\AField{\_⊢↑\_}}.
  \item
    The lambda typing constructor \ACode{\ACon{⊢λ}} weakens the type \ACode{t₂} to
    shield it from expression-substitution, requiring us to use the
    \ACode{\AField{⋯-↑-wk}} lemma.
  \item
    The type application constructor \ACode{\ACon{⊢∙}} substitutes \ACode{t₁} into \ACode{t₂},
    requiring us to use \ACode{\AField{dist-↑-⦅⦆}}.
  \end{itemize}
  % Finally, we instantiate the \ACode{\ACon{TTraversal}} structure:
  % \AppCode\FTTraversal

@@

  \section{Reflection \& Generics}
  \label{sec:reflection}
  We used Agda's reflection mechanism to derive
  instantiations related to all structures for untyped substitition,
  i.e.\ \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
  \ACode{\ACon{CTraversal}}.

  To gain insight into the class of object languages supported by our
  reflection algorithm, we have formalized a class of object languages
  using a generic programming technique and instantiated the structures
  generically. Informally, this class contains all object languages described
  via multi-sorted syntax, which
  \begin{enumerate}
  \item have a variable constructor of type
    \ACode{S \ACon{∋} s \AKw{→} S \ACon{⊢} s};
  \item use subterms only directly (e.g. not in lists);
  \item only extend the scope-context of subterms, but never modify it.
  \end{enumerate}

  Our reflection algorithm derives proofs with exactly the same
  structure as the generic proof, giving high confidence that it covers
  the same class of languages.

  The formal definition of the generic class can be found in
  the supplementary material.
  % Appendix~\ref{sec:generic-class} and the actual implementation.

  We believe restriction 2 is purely technical and can be lifted
  with a more sophisticated reflection algorithm, as parameterized
  datastructures can also be inlined into the syntax definition as terms
  of a new sort.

>>>>

  \AppCode\FPreserve
  The type of \ACode{\ADef{\_⊢⋯\_}} is the same as in the record definition and hence omitted.
  The interesting parts of the proof are:
  \begin{itemize}
  \item
    There is a strong similarity to the instantiation of
    map traversal \ACode{\AField{\_⋯\_}}. Where \ACode{\AField{\_⋯\_}}
    used \ACode{\AField{`/id}} or \ACode{\AField{\_↑\_}}, we now use their
    preservation lemmas \ACode{\AField{⊢`/id}} or \ACode{\AField{\_⊢↑\_}}.
  \item
    The lambda typing constructor \ACode{\ACon{⊢λ}} weakens the type \ACode{t₂} to
    shield it from expression-substitution, requiring us to use
    \ACode{\AField{⋯-↑-wk}} to move the map under the weakening.
  \item
    The type application constructor \ACode{\ACon{⊢∙}} substitutes \ACode{t₁} into \ACode{t₂},
    requiring us to use \ACode{\AField{dist-↑-⦅⦆}} to move the map
    under the singleton substitution.
  \end{itemize}
  % Finally, we instantiate the \ACode{\ACon{TTraversal}} structure:
  % \AppCode\FTTraversal

@@

  \section{Reflection \& Generics}
  \label{sec:reflection}
  We use Agda's reflection mechanism to derive
  instantiations related to all structures for untyped substitition,
  i.e.\ \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
  \ACode{\ACon{CTraversal}}.

  To gain insight into the class of object languages supported by our
  reflection algorithm, we have instantiated the structures for a generic syntax
  similar to the one in Allais et al.\cite{DBLP:journals/pacmpl/AllaisA0MM18}.
  Our reflection algorithm derives proofs with the same
  structure as the generic proofs, giving high confidence that it covers
  the same class of languages.

  Informally, all objects languages with multi-sorted syntax are supported that
  \begin{enumerate}
  \item have a variable constructor of type
    \ACode{\AKw{∀} \{S s\} \AKw{→} S \ACon{∋} s \AKw{→} S \ACon{⊢} s};
  \item use subterms only directly (e.g. not in lists); and
  \item only extend the scope-context of subterms, but never modify it otherwise.
  \end{enumerate}

  % Appendix~\ref{sec:generic-class} and the actual implementation.

  We believe restriction 2 is purely technical and can be lifted
  with a more sophisticated reflection algorithm. In the current system,
  this restriction can be worked around by inlining the data structure
  constructors into the syntax definition as terms of a new sort.

  The formal definition of the generic syntax can be found in
  the supplementary material.
